---
title: "STATS769 Lab1"
author: "Yujie Zhang 130011770 yzhb915"
date: "August 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. The dataset is of csv format and was imported by following steps:

```{r}
## Read the data from csv files
trips_2018_7 <- read.csv("./trips-2018-7.csv", header = TRUE) 
trips_2018_8 <- read.csv("./trips-2018-8.csv", header = TRUE) 
trips_2018_9 <- read.csv("./trips-2018-9.csv", header = TRUE)
```

2. The training and test sets are created as following:

```{r}
## Combine into single object
trips <- rbind(trips_2018_7, trips_2018_8, trips_2018_9)
## Select 1000 samples from each month as test dataset
trips_test_201807 <- sample(1:5000,1000)
trips_test_201808 <- sample(5001:10000,1000)
trips_test_201809 <- sample(10001:15000,1000)
trips_test_index <- rbind(trips_test_201807, trips_test_201808, trips_test_201809) 
trips_test <- trips[trips_test_index,]
trips <- trips[-trips_test_index,]
```

3. Data exploration and summary of variables:

```{r}
## Distribution of duration
plot(density(trips$duration))
plot(density(log(trips$duration)))
```

The distribution of duration is more like a normal distribution after the log transformation.

```{r}
## Relationship between distance and duration times
plot(trips$distance, trips$duration)
plot(trips$distance, trips$duration, log="xy")

## Check missing data 
library(visdat) 
vis_miss(trips)
```

4. Model fitting using a training dataset:

```{r}
## Transform training dataset
trainDuration <- log(trips$duration) 
trainSubset_duration <- is.finite(trainDuration)
trainDistance <- log(trips$distance) 
trainSubset_distance <- is.finite(trainDistance)
trainSubset <- trainSubset_duration & trainSubset_distance
## Linear Model with training dataset
yTrain <- trainDuration[trainSubset]
xTrain <- trainDistance[trainSubset]
fitMean <- mean(yTrain, na.rm=TRUE)
fitLM <- lm(y ~ x, data.frame(y=yTrain, x=xTrain))
## Visualize the model
plot(xTrain, yTrain) 
abline(h=fitMean, col="red", lwd=3) 
abline(fitLM, col="blue", lwd=3)

## Transform test dataset
testDuration <- log(trips_test$duration) 
testSubset_duration <- is.finite(testDuration) 
testDistance <- log(trips_test$distance) 
testSubset_distance <- is.finite(testDistance) 
testSubset<- testSubset_duration&testSubset_distance
## Linear Model with testing dataset
yTest <- testDuration[testSubset]
xTest <- testDistance[testSubset]
predMean <- rep(fitMean, length(yTest)) 
predLM <- predict(fitLM, data.frame(x=xTest))
```

5. Model evaluation using a test dataset:

```{r}
## Evaluate models
RMSE <- function(m, o) 
{
  sqrt(mean((m - o)^2)) 
}
RMSE(predMean, yTest)
```

```{r}
## Visualise models
RMSE(predLM, yTest)
plot(xTest, yTest) 
abline(h=predMean, col="red", lwd=3) 
abline(fitLM, col="blue", lwd=3)
```

6. Conclusion: According to the value of RMSE, it fits linear mode well and the prediction is good.
