---
title: "STATS769 Lab02"
author: "Yujie Zhang 130011770 yzhb915"
date: "August 11, 2019"
output: pdf_document
---

The data are trips on electric scooters and bikes in Austin, Texas.The data came in the form of three CSV files, one per month.  Each file contains 5000 trips.The following code reads the CSV files into R and combines them to create a single data frame.

1.The following code imported the eleven CSV files into R and combine them into a single data frame.
```{r}
file_path="/course/Labs/Lab02"
myfiles = list.files(path=file_path, pattern="trips-20", full.names=TRUE)
myfiles
trips <- do.call(rbind, lapply(myfiles, read.csv))
```

2.A test set was generated by selecting 1000 rows at random from each month. The following code extracted a subset of 1000 rows from each month to use as a test set (a total of 11,000 rows); the remaining 44,000 rows are the training set.

```{r}
testIndex <- unlist(lapply(1:11,function(i) {sample(1:5000, 1000) + (i - 1)*5000}))
test_data = trips[testIndex,]
train_data = trips[-testIndex,]
```

Exploratory plots of trip durations and trip distances shows they are reasonably unimodal and symmetric.


3.The following code transformed the training set by removing trips with a distance or duration that is non-positive, then log the duration variable. And then we plot the density of train duration.Exploratory plots of trip durations shows it is reasonably unimodal and symmetric.

```{r}
trainIndex <- train_data$Trip.Duration>0 & train_data$Trip.Distance>0
trainsubset <- train_data[trainIndex,]
trainDuration <- log(trainsubset$Trip.Duration)
trainDistance <- trainsubset$Trip.Distance
plot(density(trainDuration),main="Distribution of Trip Durations\n(training set)")
```


The following plot shows the relationship between duration and distance. The first plot indicates that there is outliers in the data. After trying with log transformation, it looks better.

```{r}
plot(trainDuration ~ trainDistance)
plot(trainDuration ~ log(trainDistance))
```


4.To predict the probobility of long trip, we need to transform the data of train distance to be a categorial value. The following code created a “long trip” variable that is TRUE if the trip distance is greater than 1000 (1km) and FALSE otherwise. And then show the relationship between long trip and train duration.

```{r}
train_data$`long trip` <- train_data$Trip.Distance>1000
trainLongtrip <- train_data$`long trip`[trainIndex]
boxplot(trainDuration ~ trainLongtrip, col = c("green","red"))
```


5.The following code fitted a logistic regression model to predict the proportion of long trips based on trip duration using training dataset.

```{r}
fitGLM <- glm(y ~ x, data.frame(x=trainDuration, y=trainLongtrip), family="binomial",na.action=na.exclude)
summary(fitGLM)
```


6.Before the prediction, we prepared and transformed the test set in the same way as we transformed the training set.

```{r}
testIndex <- test_data$Trip.Duration>0 & test_data$Trip.Distance>0
testsubset <- test_data[testIndex,]
testDuration <- log(testsubset$Trip.Duration)
testDistance <- testsubset$Trip.Distance

test_data$`long trip` <- test_data$Trip.Distance>1000
testLongtrip <- test_data$`long trip`[testIndex]
```

7.We did the prediction and then visualized and evaluated the model on the test set.

```{r}
predGLM <- predict(fitGLM, data.frame(x=testDuration), type="response")
predProp <- mean(testLongtrip, na.rm=TRUE)
breaks = seq(1,10,1)
testBlocks <- cut(testDuration, breaks=breaks)
testProps <- tapply(testLongtrip, testBlocks, mean)
testProps
plot(seq(1, 9, 1), testProps)
abline(h=.5, lty="dashed")
abline(h=predProp, col="red", lwd=3)
o <- order(testDuration)
lines(testDuration[o], predGLM[o])

table(testLongtrip, rep(predProp > .5, length(testLongtrip)))
table(testLongtrip, predGLM > .5)
library(caret)
confusionMatrix(factor(rep(predProp > .5, length(testLongtrip))),
                factor(testLongtrip))
confusionMatrix(factor(predGLM > .5),
                factor(testLongtrip))
## Vary threshold
confusionMatrix(factor(predGLM > .4),
                factor(testLongtrip))
## Guessing model
confusionMatrix(factor(sample(c(TRUE, FALSE), length(testDistance), replace=TRUE)),
                factor(testLongtrip))
```

## Conclusion
The data for this exercise came in a CSV format, which presented no difficulties for importing into R. Generating a test set was slightly complicated by requiring sub-samples from each month, though this only required standard R data manipulation tools. 

We required a log-transformation of duration variables to obtain unimodal, symmetric data for model fitting.  This required excluding non-positive distances and durations (about 8% of the data),which means that our predictive model is only valid for trips that had a non-zero distance or duration. We are unlikely to be predicting zero-distance or zero-duration trips, so this is not a large concern.

We fitted a logistic regression model, which provided some predictive power, though there are clear opportunities for a more flexible model(or a more robust model) to perform better.

According to the result of evaluation, the accuracy, sensitivity and specificity are all about 0.82 respectively with the threshold of 0.5. The details of numbers are displayed in the above result. So the model is good for the analyze. We also tried to change the threshold to be 0.4, and there is little change for the accuracy, sensitivity and specificity. 

We finished this experiment on linux virtual machine with commands like ssh, scp,ls,cd,grep,vim,R. The difference is is that, since there is no desktop of the virtual machine, we can not use RStudio. 
We need to generate the report by using linux command: 

Rscript -e "rmarkdown::render('./STATS769_Lab02_YujieZhang.rmd')"

Since linux system is widely used in different servers in the environment of production, it's still convinent to process big files on the servers.
